Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-04-15 article.tex
Add Mycoplasma

For MHC-I, it was found that predicted epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all HLA-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs have a reduced variability 
due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.

@@

If presentation of TMHs on MHC-I would bring an evolutionary advantage 
in the recognition of pathogens by the immune system, 
it would follow that this is equally important for MHC-II, 
especially as CD4+ T cell help is needed for a long lasting CD8+ T cell 
response \cite{novy2007cd4}.

@@

\paragraph{Evolutionary signal of unknown strength}

\richel{Check against stupidity}
We do not know how viruses are selected \richel{so I should read the
literature on that}. We predict that there is a strong 
evolutionary conservation in regions that are essential for viral proteins'
functions. Because these regions are rare in a viral proteome,
and differ widely between pathogens, we expect that the human immune system
is not tailored to detect these sites. TMHs, however, have less 
functional constraints yet are more general. Because TMHs are more general,
we expect it for the immune system to pay off to be more attentive towards
TMHs. Because the strength of the evolutionary signal,
are strong data set needs to be used.

We expect that

We can assume that only after the moment that COVID-19 spilled over 
to humans, it's the human immune system that is an influential component 
of its evolutionary selection.
Using the extensive amount of time-dated COVID-19 sequences,
we can quantify to which extent selection in COVID-19 occurs
in the transmembrane part of its proteome.

@@

\begin{itemize}
  \item $\mathcal{H}_1$: MHC-I has the same percentage of epitopes overlapping
    with TMHs in Homo sapiens as in COVID-19
  \item $\mathcal{H}_2$: MHC-II has the same percentage of epitopes overlapping
    with TMHs as expected by chance
  \item $\mathcal{H}_3$: The TMHs of COVID-19 have the same mutation rate 
    as the cytosolic and extracellular parts of its transmembrane proteins
\end{itemize}

@@

We used the same analysis as \cite{bianchi2017},
except that instead of a human reference proteome,
we used the proteome of the first sequenced COVID-19 strain (\cite{wu2020new}),
with a GenBank ID of MN908947.3, as can be downloaded 
from \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947}.

Bianchi and colleagues obtained a distribution of 
percentages of MHC-I epitopes overlap with TMHs in Homo sapiens
for the different HLA haplotypes, with an average of 5.3\%.
We obtained a similar distribution of percentages of MHC-I epitopes that 
overlap with TMHs for the different HLA haplotypes, but then applied to
COVID-19.

We compare these distributions using a two-sample 
Kolmogorov-Smirnov (KS) test. The KS test determines if two samples
are derived from the same distribution, without making assumptions
regarding the shape of that distribution.

We will reject
the null hypothesis that MHC-I has the same percentage of epitopes 
overlapping with TMHs in Homo sapiens as in COVID-19 when 
the KS statistic $D_{n,m}$ follows the relationship as shown in 
equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
and $n = m = 13$ HLA haplotypes.

@@

Figure \ref{fig:1} shows the percentages of epitopes overlapping 
with TMHs for a human and COVID-19 proteome.
The KS test to determine if these percentages are sampled from
a same distribution resulted in a p-value of \richel{unknown},
which makes us reject/accept the hypothesis that these percentages
are sampled from the same distribution.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1/fig_bbbq_1_watermarked.png}
  \caption{
    \% epitopes overlapping with transmembrane helix
    for a human and COVID-19 proteome.
    \richel{
      The Illuminati watermark indicates the data underlying this
      figure has been simulated
    }
  }
  \label{fig:1}
\end{figure}

@@

@@

>>>>

\richel{Check against stupidity}
For MHC-I, it was found that predicted epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all HLA-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs have a reduced variability 
due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.
Note that the mechanism by which a cell presents its TMHs is
yet unknown.

@@

\richel{Check against stupidity}
If presentation of TMHs on MHC-I would bring an evolutionary advantage 
in the recognition of pathogens by the immune system, 
it would follow that this is equally important for MHC-II, 
especially as CD4+ T cell help is needed for a long lasting CD8+ T cell 
response \cite{novy2007cd4}. 
The mechanism to detect foreign TMHs by MHC-II would be unknown, 
similar to the discovery that MHC-I presents TMHs.

@@

\paragraph{Evolutionary signal of unknown strength for virus}

\richel{Check against stupidity}
Regarding human viruses, we do not know the extent to which the human
immune systems selects on these, nor which loci are selected
upon \richel{so I should read the literature on that}. 
We predict that there is a strong 
evolutionary conservation in regions that are essential for viral proteins'
functions. Because these regions are rare in a viral proteome,
and differ widely between pathogens, we expect that the human immune system
is not tailored to detect these sites. TMHs, however, have less 
functional constraints yet are more general. Because TMHs are more general,
we expect it for the immune system to pay off to be more attentive towards
TMHs. 
Additionally, viruses are not only selected for their evasion of an 
immune response, yet are selected to have a high reproductive value.
Because the strength of the evolutionary signal is unknown,
we use a strong data set to be able to detect it.

\paragraph{Evolutionary signal of unknown strength for bacteria}

\richel{Check against stupidity}
Regarding human pathogenic bacteria, we do not know the extent to which the 
human immune systems selects on these, nor which loci are selected
upon \richel{so I should read the literature on that}.
Because most bacterial pathogens are generalists, that is,
can infect multiple hosts, we expect the selection pressure exerted
by the human immune response to be weak.
Also, it is unknown if bacterial TMHs are detected by MHC-II at all.
Due to this, the strength of the evolutionary signal is unknown,
thus we need a strong data set to be able to detect it.

\paragraph{Data sets}

\richel{Check against stupidity}
To detect an evolutionary signal for selection on the TMHs of viruses, 
we use the many time-dated COVID-19 proteomes.
We can assume that only after the moment that COVID-19 spilled over 
to humans, it has mostly been the human immune system selecting against
its detection, although this is only part of the selection for 
having a high transmission rate.
To detect an evolutionary signal for selection on the TMHs of bacteria, 
we use the proteomes of Mycoplasma. Mycoplasma is \richel{something}.

@@

\begin{itemize}
  \item $\mathcal{H}_1$: MHC-I has the same percentage of epitopes overlapping
    with TMHs in Homo sapiens as in COVID-19 as in Mycoplasma
  \item $\mathcal{H}_2$: MHC-II has the same percentage of epitopes overlapping
    with TMHs as expected by chance in Homo sapiens and COVID-19 and Mycoplasma
  \item $\mathcal{H}_3$: The TMHs of both COVID-19 and Mycoplasma
    have the same mutation rate 
    as the cytosolic and extracellular parts of their transmembrane proteins
\end{itemize}

@@

We used the same analysis as \cite{bianchi2017},
except that instead of a human reference proteome,
we use the proteome of the first sequenced COVID-19 strain (\cite{wu2020new},
GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
as well as a Mycoplasma \richel{that one} reference genome.

Bianchi and colleagues obtained a distribution of 
percentages of MHC-I epitopes overlap with TMHs in Homo sapiens
for the different HLA haplotypes, with an average of 5.3\%.
We obtained a similar distribution of percentages of MHC-I epitopes that 
overlap with TMHs for the different HLA haplotypes, but then applied to
COVID-19 and Mycoplasma.

We compare the distributions of humans and each pathogen
using a two-sample 
Kolmogorov-Smirnov (KS) test. The KS test determines if two samples
are derived from the same distribution, without making assumptions
regarding the shape of that distribution.

We will reject
the null hypothesis that MHC-I has the same percentage of epitopes 
overlapping with TMHs in Homo sapiens compared to each pathogen when 
the KS statistic $D_{n,m}$ follows the relationship as shown in 
equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
and $n = m = 13$ HLA haplotypes.

@@

Figure \ref{fig:1} shows the percentages of epitopes overlapping 
with TMHs for a human, COVID-19 and Mycoplasma proteome.
The KS test to determine if these percentages are sampled from
a same distribution resulted in a p-value of \richel{unknown},
which makes us reject/accept the hypothesis that these percentages
are sampled from the same distribution.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1/fig_bbbq_1_watermarked.png}
  \caption{
    \% epitopes overlapping with transmembrane helix
    for a human, COVID-19 and Mycoplasma proteome.
    \richel{
      the data underlying this figure has been simulated
    }
  }
  \label{fig:1}
\end{figure}

@@

We found that the percentages of epitopes overlapping 
with TMHs for a human and Mycopasma proteome are 
[similar/different]. In other words, the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its bacterial pathogen.

@@

We concluded that the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its bacterial pathogen.
Because a bacterium does not infect a cell, thus its polypeptides
will not be presented by MHC-I, this result is [unexpected/expected]
